Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data

### 2022-11-11 paper/manuscript.Rmd
LC comment; source code

Byron C. Jaeger, PhD,^1^ 
Ligong Chen, PhD,^2^
Kathryn Foti,^2^
Shakia T. Hardy,^2^
Adam P. Bress, PharmD, MS,^3,4^ 
Sean P. Kane, PharmD, BCPS,^5^
Jennifer S. Herrick, MS,^3,6^
Catherine Derington, PharmD, MS,^7^ 
Bharat Poudel, MSPH,^2^
Ashley Christenson, BS,^2^ 
Lisandro D. Colantonio, PhD,^2^
Paul Muntner, PhD,^2^

@@

^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
^2^Department of Epidemiology, University of Alabama at Birmingham. `r single_par`
^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT. `r single_par`
^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
^5^Department of Pharmacy Practice, Rosalind Franklin University of Medicine and Science, North Chicago, IL, USA
^6^Department of Internal Medicine, University of Utah School of Medicine, Salt Lake City, UT
^7^ University of Colorado Skaggs School of Pharmacy and Pharmaceutical Sciences, Aurora, CO, USA
\newpage

@@

**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming. The application’s interface allows users to estimate crude and age-adjusted means, quantiles, and proportions. Population counts can also be estimated. To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.The application was validated by replicating statistics from three publications.

**Results**: The estimated mean systolic blood pressure (BP) declined from 123 mmHg in 1999-2000 to 120 mmHg in 2009-2010 and increased to 123 mmHg in 2017-2020. The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was 47.9% in 1999-2000, 43.0% in 2009-2010, and 44.7% in 2017-2020. In 2017-2020, 115.3 million US adults had hypertension. The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020. After age-adjustment and among US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.

@@

We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming. The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible. In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.

>>>>

Byron C. Jaeger, PhD,^1^ 
Ligong Chen, PhD,^2^
Kathryn Foti,^2^
Shakia T. Hardy,^2^
Adam P. Bress, PharmD, MS,^3,4^ 
Sean P. Kane, PharmD, BCPS,^5^
Jennifer S. Herrick, MS,^3,6^
Catherine Derington, PharmD, MS,^7^ 
Bharat Poudel, MSPH,^2^
Ashley Christenson, BS,^2^ 
Lisandro D. Colantonio, MD, PhD,^2^
Paul Muntner, PhD,^2^

@@

^1^Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, NC. `r single_par`
^2^Department of Epidemiology, University of Alabama at Birmingham, Birmingham, AL. `r single_par`
^3^Informatics, Decision-Enhancement, and Analytic Sciences (IDEAS) Center, Veterans Affairs, Salt Lake City Health Care System, Salt Lake City, UT. `r single_par`
^4^Department of Population Health Sciences, University of Utah School of Medicine, Salt Lake City, UT. `r single_par`
^5^Department of Pharmacy Practice, Rosalind Franklin University of Medicine and Science, North Chicago, IL, USA
^6^Department of Internal Medicine, University of Utah School of Medicine, Salt Lake City, UT
^7^ University of Colorado Skaggs School of Pharmacy and Pharmaceutical Sciences, Aurora, CO, USA
\newpage

@@

**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020, without requiring users to conduct statistical programming. The application's interface allows users to estimate crude and age-adjusted means, quantiles, and proportions. Population counts can also be estimated. To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.The application was validated by replicating statistics from three publications.

**Results**: The estimated mean systolic blood pressure (BP) declined from 123 mmHg in 1999-2000 to 120 mmHg in 2009-2010 and increased to 123 mmHg in 2017-2020. The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was 47.9% in 1999-2000, 43.0% in 2009-2010, and 44.7% in 2017-2020. In 2017-2020, 115.3 million US adults had hypertension. The age-adjusted prevalence of controlled BP, defined by the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.7% in 1999-2000, 25.0% in 2013-2014, and 21.9% in 2017-2020. After age-adjustment and among non-pregnant US adults who self-reported taking antihypertensive medication, 27.6%, 48.5%, and 43.0% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively.

@@

We developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming. The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid and reproducible. The application is also "open-source" as it's source code is publicly available. In this manuscript, we review the design, development and validation of the application and present BP and hypertension statistics for US adults that were generated using the application.
